Log or not - Analysis of ECDC forecasts

### 2022-01-31 analysis/Analysis ECDC forecasts.Rmd
update plots

hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type, location) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ target_type, scales = "free") + 
  theme_minimal() + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
  labs(y = "Mean observed value", x = "Location")

@@

```{r}
# alternatively, we could think about just summarising across all models for these plots? 
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  labs(y = "Interval score", x = "Target type")

@@

  scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(wis = summary_fct(interval_score), 
            .groups = "drop_last") |> 
  inner_join(total_values) |>
  ggplot(aes(y = wis, x = total)) + 
  geom_point() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  # scale_x_continuous(labels = scale_fn) + 
  # scale_y_continuous(labels = scale_fn) + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") +
  labs(y = "REPLACE", x = "Total number of observed values")
}

mean_wis_vs_total <- wis_vs_total(scores)
ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)

sd_wis_vs_total <- wis_vs_total(scores, summary_fct = stats::sd)
ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)

>>>>

hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type, location) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ target_type, scales = "free") + 
  theme_minimal() + 
  scale_y_continuous(labels = label_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
  labs(y = "Mean observed value", x = "Location")

@@

```{r}
# alternatively, we could think about just summarising across all models for these plots?

# make two separate plots to keep ordering. I.e. order locations according 
# to log scale to illustrate that the ordering changes for cases

get_order <- function(df, type = "Cases") {
  df |> 
    filter(scale == "natural", target_type == type) |>
    arrange(-interval_score) |>
    pull(location)
}

plot_df <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last")

order_cases <- get_order(plot_df)
order_deaths <- get_order(plot_df, type = "Deaths")

plot <- function(plot_df, order) {
  plot_df |>
    mutate(location = factor(location, levels = order)) |>
    ggplot(aes(y = interval_score, x = location)) + 
    geom_bar(stat = "identity") + 
    scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
    facet_wrap(~ target_type + scale, scale = "free") + 
    theme_minimal() + 
    theme(legend.position = "none") + 
    scale_y_continuous(labels = scale_fn) + 
    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
    labs(y = "Interval score", x = "Location")
}

cases <- plot_df |>
  filter(target_type == "Cases") |>
  plot(order = order_cases)

deaths <- plot_df |>
  filter(target_type == "Deaths") |>
  plot(order = order_deaths)

cases / 
  deaths

@@

  scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(wis = summary_fct(interval_score), 
            .groups = "drop_last") |> 
  inner_join(total_values) |>
  ggplot(aes(y = wis, x = total)) + 
  geom_point() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  scale_x_continuous(labels = scale_fn) +
  scale_y_continuous(labels = scale_fn) +
  labs(y = "REPLACE", x = "Total number of observed values")
}

wis_vs_total(scores) + 
  labs(y = "Mean WIS")
ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)

wis_vs_total(scores, summary_fct = stats::sd) + 
  labs(y = "Sd WIS")
ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)

# log-log versions of that plot
wis_vs_total(scores) + 
  labs(y = "Mean WIS") + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") 
ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 5)

wis_vs_total(scores, summary_fct = stats::sd) + 
  labs(y = "Sd WIS") + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") 
ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
